Future challenges for Research Software Engineering

### 2025-04-17 future_challenges_for_RSEng.tex

@@

\subsection{Impact of Alternate Career Paths for Researchers on the Structure of Research Institutes}
\begin{whatis}{text 1 }{text 2}
FG-NOTE: By Nick. I've put it here, due to the realtion with the RTP question.
\end{whatis}

>>>>

\subsection{How have academic Institutions transformed by then? ( Point 4 in the issue )}
\begin{whatis}{What do we want to have achieved}{acedemic}
We hope to see that universities use the presence of an RSE Department actively as a means to differentiate themselves from other universities.
We hope that the central RSE units have local RSE representatives, or even small RSE groups that handle tasks within a department to oversee specific projects[FIXME: cite group paper].
We also hope, that specialized clusters form, with certain RSE units specializing in certain tasks like data management, HPC, or visualization.
At universities which have them, we want to see, that researchers naturally accept them as part of their used services.
\end{whatis}

\subsubsection{Which challenges will still persist}
The workshop participants expect that there will still be budgeting and administrative hurdles to creating
permanent and well-funded RSE teams.
While we hope that the barrier to researchers diminishes, the workshop participants
still expect some work to do. We hope that the compartmentalization that occurs will help researchers reprioritize their workload.
An upcoming challenge will be that, with the emergence of more dedicated RSE units,
networking between them will come into the foreground, because there should be a strong need to share best practices
among institutions with succesful RSE Teams.
On a more organisational level, we expect the sharing of models, to fit the different organisational constraints.

What are open questions here?
Organizations will have to come up with models for hiring people. While
a standardized RSE masters makes it easier, it will be interesting to see, if
RSE are still often hired by word of mouth or if the master's really makes it more comparable.

@@

\subsection{Impact of Alternate Career Paths for Researchers on the Structure of Research Institutes}
FG-TEXT-START: The RSE is probably one of the first professionalizations in the academic midlevel that is able
to collectively voice their concerns. This enables the possibility to reshape universities.
If more people are employed at the midlevel, this makes a steadier pool of people
available that can integrate new people that are regularly coming into departments and know the local facilities.
On the other hand an upcoming challenge can be, how universities decide to design the contracts of people employed
in RSE units. will they be permanent, or will there also be an up-or-out mentality?
FG-TEXT-END

FG-TEXT-START:
As outlined in [cite RSE group paper] not all universities need to opt for strengthening
the academic midlevel. Some might have chosen to set up external companies to outsource typical RSE tasks,
thereby providing career paths outside of academia.
FG-TEXT-END:

\begin{story}{text 1 }{text 2}
FG-NOTE: By Nick. I've put it here, due to the realtion with the RTP question.
\end{story}
